---
title: "exploratory numeric analysis"
author: "Jiayi Shi"
date: "2022-10-05"
output: github_document
---
```{r results='hide',message=FALSE}
library(tidyverse)

Sys.setlocale("LC_TIME", "English")

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


```{r message=FALSE}
weather_df =  
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  # 2012-01-13 rounded to minth: 2012-01-01
  select(name, id, everything())
```

## `group_by`+`summarise()`

`group_by`: don't change any of the data, just add a layer on top of the data set, change the behavior of functions used  it.

```{r}
weather_df %>% 
  group_by(name, month) 
```


### Counting things

* `group_by`+`summarise(n())`
```{r}
weather_df %>% 
  group_by(name, month) %>% # can group by more than one variable
  summarise(n_obs = n())
```

* `count()`
```{r}
weather_df %>% 
  count(name,month,name = "n_obs")#the count column name, default to be "n"
```

* don't use `table`
```{r}
weather_df %>%
  pull(month) %>% 
  table
```

* multiple summaries
```{r}
weather_df %>%
  group_by(month) %>%
  summarize(
    n_obs = n(),
    n_days = n_distinct(date))
```

### 2x2 tables

Tabulate the frequency of a binary outcome across levels of a binary predictor.

* ``group_by` + `summarise` + `pivot_wider` 

```{r}
weather_df %>%
  mutate(
    cold = case_when(
      tmax>=5 ~ "not_cold",# condition ~ output value
      tmax<5 ~ "cold",
      TRUE ~ "" # force case_when to output value behind TRUE if none of the previous conditions were TRUE
    ) 
  ) %>% 
  filter(name != "Waikiki_HA") %>% 
  group_by(name,cold) %>% 
  summarise(count = n()) %>% 
  pivot_wider(names_from = "cold",values_from = "count")
 # convert to a nice 2x2 table
```

* `knitr::ktable()`: table for Rmarkdown file -- Presenting reader-friendly results for this kind of exploratory analysis. (use pivot_wider to untidy data so that reader-friendly)
```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(mean_tmax = mean(tmax)) %>% 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax) %>% 
  knitr::kable(digits = 1) # digits: decimal place of output values in the table
```


* janitor::tabyl (generate frequency table)

```{r}
weather_df %>% 
  mutate(cold = case_when(
    tmax <  5 ~ "cold",
    tmax >= 5 ~ "not_cold",
    TRUE     ~ ""
  )) %>% 
  filter(name != "Waikiki_HA") %>% 
  janitor::tabyl(name, cold)
```

### functions in `summaries()`

```{r}
weather_df %>%
  group_by(month) %>%
  summarize(
    mean_tmax = mean(tmax),
    mean_prec = mean(prcp, na.rm = TRUE),
    median_tmax = median(tmax),
    sd_tmax = sd(tmax))
```

* `across()`

Used to summarize multiple columns using the same summary.

```{r}
weather_df %>%
  #group_by(name, month) %>%
  summarize(across(tmin:prcp, mean, na.rm=T))
```

### Plot

```{r}
weather_df %>% 
  group_by(name,month) %>% 
  summarise(mean_tmax = mean(tmax)) %>% 
  # prepare the dataframe before plotting
  
  ggplot(aes(x = month, y = mean_tmax, color = name)) + 
  geom_point() +
  geom_path()
```

## `group_by`+`mutate()`

To compare the daily max temperature to the annual average max temperature for each station separately and plot the results.
```{r}
weather_df %>%
  group_by(name) %>%
  mutate(
    mean_tmax = mean(tmax, na.rm = TRUE),
    centered_tmax = tmax - mean_tmax) %>% 
  ggplot(aes(x = date, y = centered_tmax, color = name)) + 
    geom_point() 
```

### Window functions
 aggregate funcitons: n inputs, 1 output  
 window functions: n inputs, n outputs

* `min_rank()`

```{r}
weather_df %>%
  group_by(name, month) %>%
  mutate(temp_ranking = min_rank(tmax)) # min_rank(): one way of ranking 
```
 
Useful when filtering data based on rank:

```{r}
weather_df %>%
  group_by(name, month) %>%
  filter(min_rank(tmax) < 2)
# keep only the day with the lowest max temperature within each month
```

```{r}
weather_df %>%
  group_by(name, month) %>%
  filter(min_rank(desc(tmax)) < 4)
# keep the three days with the highest max temperature
```
 
* `lag()`
used to compare an observation to it’s previous value:

```{r}
weather_df %>%
  group_by(name) %>%
  mutate(
    yesterday_tmax = lag(tmax) # default lag one day; should group_by first
  )
```


```{r}
weather_df %>%
  group_by(name) %>%
  mutate(temp_change = tmax - lag(tmax))
# find the day-by-day change in max temperature within each station over the year
```

This kind of variable might be used to quantify the day-by-day variability in max temperature, or to identify the largest one-day increase:

```{r}
weather_df %>%
  group_by(name) %>%
  mutate(temp_change = tmax - lag(tmax)) %>%
  summarize(
    temp_change_sd = sd(temp_change, na.rm = TRUE),
    temp_change_max = max(temp_change, na.rm = TRUE))
```

## Assessment

```{r}
pulse = haven::read_sas("data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>% 
  pivot_longer(
    4:7,
    names_to = "visit",
    names_prefix = "bdi_score_",
    values_to = "bdi",
  ) %>% 
  select(id,visit,everything()) %>% 
  mutate(
    visit = recode(visit,"bl"="00m"),
    visit = factor(visit,levels = str_c(c("00","01","06","12"),"m"))
    #visit = as.factor(visit)
  ) %>% 
  arrange(id,visit)

pulse %>% 
  group_by(visit) %>% 
  summarise(mean_bdi = mean(bdi, na.rm = T),
            median_bdi = median(bdi, na.rm = T)) %>% 
  knitr::kable(digits = 2)
```


```{r}
pup_data = 
  read_csv("./data/FAS_pups.csv") %>%
  janitor::clean_names() %>%
  mutate(sex = recode(sex, `1` = "male", `2` = "female")) 

litter_data = 
  read_csv("./data/FAS_litters.csv") %>%
  janitor::clean_names() %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3)

fas_data = left_join(pup_data, litter_data, by = "litter_number")

fas_data %>% 
  group_by(dose, day_of_tx) %>% 
  drop_na(dose) %>% 
  summarize(mean_pivot = mean(pd_pivot, na.rm = TRUE)) %>% 
  pivot_wider(
    names_from = dose, 
    values_from = mean_pivot) %>% 
  knitr::kable(digits = 3)
```


